Parallel Polychoric Test Suite
===============================

This is a series of tests for the parallelized `polychoric()` function.  The speedup 
is acheived using the [foreach](http://cran.r-project.org/web/packages/foreach/index.html) 
and [doMC](http://cran.r-project.org/web/packages/doMC/index.html) packages. 

Using **doMC**, we register the backend by telling R that we have *n* cpus available to use
instead of one.  Then, we use **foreach** as the frontend, and replace the `for` 
loop in `psych:::polyBinBvn` with `foreach(...) %dopar% {}`.  Instead of iterating 
over each individual cell to figure out the optimal correlation value, we can do *n* 
at a time.

Simulate our data
-----------------
```{r cache=TRUE,tidy=TRUE,highlight=TRUE}
library(psych)
poly.data <- sim.irt(nvar=50,n=100)
```

Bill's Speed
-------------
```{r cache=TRUE,tidy=TRUE,highlight=TRUE}
system.time(bill <- psych:::polychoric(x=poly.data$items))
```

Bill's Output
------------
```{r cache=TRUE,tidy=TRUE,highlight=TRUE}
print(bill)
```

New R Functions
--------------
```{r cache=TRUE,tidy=TRUE,highlight=TRUE}
polyBinBvn <- function (rho, rc, cc) 
{
    if (min(rc) < -9999) 
        rc <- rc[-1]
    if (min(cc) < -9999) 
        cc <- cc[-1]
    if (max(rc) > 9999) 
        rc <- rc[-length(rc)]
    if (max(cc) > 99999) 
        cc <- cc[-length(cc)]
    row.cuts <- c(-Inf, rc, Inf)
    col.cuts <- c(-Inf, cc, Inf)
    nr <- length(rc) + 1
    nc <- length(cc) + 1
    P <- matrix(0, nr, nc)
    R <- matrix(c(1, rho, rho, 1), 2, 2)
    for (i in 1:(nr - 1)) {
        foreach(j=1:(nc-1)) %dopar% {
            #for (j in 1:(nc - 1)) {
            P[i, j] <- pmvnorm(lower = c(row.cuts[i], col.cuts[j]), 
                               upper = c(row.cuts[i + 1], col.cuts[j + 1]), 
                               corr = R)
        }
    }
    P[1, nc] <- pnorm(rc[1]) - sum(P[1, 1:(nc - 1)])
    P[nr, 1] <- pnorm(cc[1]) - sum(P[1:(nr - 1), 1])
    if (nr > 2) {
        for (i in (2:(nr - 1))) {
            P[i, nc] <- pnorm(rc[i]) - pnorm(rc[i - 1]) - sum(P[i, 
                                                                1:(nc - 1)])
        }
    }
    if (nc > 2) {
        for (j in (2:(nc - 1))) {
            P[nr, j] <- pnorm(cc[j]) - pnorm(cc[j - 1]) - sum(P[1:(nr - 
                                                                       1), j])
        }
    }
    if (nc > 1) 
        P[nr, nc] <- 1 - pnorm(rc[nr - 1]) - sum(P[nr, 1:(nc - 
                                                              1)])
    P
}

polychoric <- function (x, smooth = TRUE, global = TRUE, polycor = FALSE, ML = FALSE, 
   std.err = FALSE, progress = TRUE, parallel = TRUE, n.cpu = NULL)
{
   if (!require(mvtnorm)) {
       stop("I am sorry, you must have mvtnorm installed to use polychoric")
   }
   if (polycor && (!require(polycor))) {
       warning("I am sorry, you must have  polycor installed to use polychoric with the polycor option")
       polycor <- FALSE
   }
   if (isTRUE(parallel)) {
       if (!require(doMC))  {
        stop('parallel not used.  Please install doMC!')}
      if (!require(foreach))  {
        stop('parallel not used.  Please install foreach!')}
      if(is.null(n.cpu)) {
        # Adapted from multicore to reduce dependencies
        # Urbanek, S. (2013)
        systems <- list(darwin = "/usr/sbin/sysctl -n hw.ncpu 2>/dev/null", 
                        freebsd = "/sbin/sysctl -n hw.ncpu 2>/dev/null", linux = "grep processor /proc/cpuinfo 2>/dev/null|wc -l", 
                        irix = c("hinv |grep Processors|sed 's: .*::'", "hinv|grep '^Processor '|wc -l"), 
                        solaris = "/usr/sbin/psrinfo -v|grep 'Status of.*processor'|wc -l")
        
        for (i in seq(systems)) if (length(grep(paste("^", names(systems)[i], sep = ""), R.version$os))) 
          for (cmd in systems[i]) {
            a <- gsub("^ +", "", system(cmd, TRUE)[1])
            if (length(grep("^[1-9]", a))) 
              n.cpu <- as.integer(a)
          }
      }
      registerDoMC(cores=n.cpu)
   }
   cl <- match.call()
   nvar <- dim(x)[2]
   nsub <- dim(x)[1]
   x <- as.matrix(x)
   xt <- table(x)
   nvalues <- max(x, na.rm = TRUE) - min(x, na.rm = TRUE) + 
       1
   if (nvalues > 8) 
       stop("You have more than 8 categories for your items, polychoric is probably not needed")
   xmin <- apply(x, 2, function(x) min(x, na.rm = TRUE))
   x <- t(t(x) - xmin + 1)
   xfreq <- apply(x, 2, tabulate, nbins = nvalues)
   n.obs <- colSums(xfreq)
   xfreq <- t(t(xfreq)/n.obs)
   tau <- qnorm(apply(xfreq, 2, cumsum))[1:(nvalues - 1), ]
   if (!is.matrix(tau)) 
       tau <- matrix(tau, ncol = nvar)
   rownames(tau) <- names(xt)[1:(nvalues - 1)]
   colnames(tau) <- colnames(x)
   mat <- matrix(0, nvar, nvar)
   colnames(mat) <- rownames(mat) <- colnames(x)
   x <- x - min(x, na.rm = TRUE) + 1
   #foreach(i=2:nvar) %dopar% {
   for (i in 2:nvar) {
       if (progress) 
           progressBar(i^2/2, nvar^2/2, "Polychoric")
       for (j in 1:(i - 1)) {
             if (t(!is.na(x[, i])) %*% (!is.na(x[, j])) > 2) {
       	        if (!polycor) {
       	          poly <- polyc(x[, i], x[, j], tau[, i], tau[, 
       	            j], global = global)
       	          mat[i, j] <- mat[j, i] <- poly$rho
       	        }
       	        else {
       	          poly <- polychor(x[, i], x[, j], ML = ML, std.err = std.err)
       	          mat[i, j] <- mat[j, i] <- poly
       	        }
       	    }
       	    else {
       	        mat[i, j] <- mat[j, i] <- NA
       	    }
       	}
       }
   diag(mat) <- 1
   if (any(is.na(mat))) {
       warning("some correlations are missing, smoothing turned off")
       smooth <- FALSE
   }
   if (smooth) {
       mat <- cor.smooth(mat)
   }
   tau <- t(tau)
   result <- list(rho = mat, tau = tau, n.obs = nsub, Call = cl)
   if (progress) {
       cat("\n")
       flush(stdout())
   }
   class(result) <- c("psych", "poly")
   return(result)
}

# Load the hidden psych functions
polyc <- psych:::polyc
polyF <- psych:::polyF
```

Parallel Speed
-------------
```{r cache=TRUE,tidy=TRUE,highlight=TRUE}
system.time(new <- polychoric(x=poly.data$items,progress=FALSE,parallel=TRUE,n.cpu=2))
```

Parallel Output
---------------
```{r cache=TRUE,tidy=TRUE,highlight=TRUE}
print(new)
```


